---
title: "Plants by Treatment"
author: "Ellen Bledsoe"
date: "10/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(portalr)
library(tidyverse)
library(vegan)

cbbPalette <- c("#000000", "#009E73", "#e79f00", "#9ad0f3", "#0072B2", "#D55E00", 
                "#CC79A7", "#F0E442")
theme_set(theme_bw())
```

## Annuals by Sampling Period and Treatment

Running CCA (from Supp et al., 2012 and Portal-LTREB) to see if there are significant differences in plant communities between controls and KR exclosures for each season of diet sampling.

``` {r annuals_data, echo=FALSE, include=FALSE, cache = TRUE}
summerannuals_2016 <- portalr::plant_abundance(type = 'Summer Annuals',
                                               level = 'plot',
                                               shape = 'crosstab') %>% 
  filter(year == '2016',
         season == 'summer',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21)) 
summerannuals_2016[, colSums(summerannuals_2016 != 0) > 0]


winterannuals_2017 <- portalr::plant_abundance(type = 'Winter Annuals',
                                               level = 'plot',
                                               shape = 'crosstab') %>% 
  filter(year == '2017',
         season == 'winter',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21))
winterannuals_2017[, colSums(winterannuals_2017 != 0) > 0]

summerannuals_2017 <- portalr::plant_abundance(type = 'Summer Annuals',
                                               level = 'plot',
                                               shape = 'crosstab') %>% 
  filter(year == '2017',
         season == 'summer',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21))
summerannuals_2017[, colSums(summerannuals_2017 != 0) > 0]

# summer annuals since plot switch
summerannuals <- portalr::plant_abundance(type = 'Summer Annuals',
                                          level = 'plot',
                                          shape = 'crosstab') %>% 
  filter(year > 2014,
         season == 'summer',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21)) 

# winter annuals since plots switch
winterannuals <- portalr::plant_abundance(type = 'Winter Annuals',
                                          level = 'plot',
                                          shape = 'crosstab') %>% 
  filter(year > 2014,
         season == 'winter',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21))
winterannuals[, colSums(winterannuals != 0) > 0]
```

```{r annuals_pcca_prep, echo=FALSE, include=FALSE}
# sqrt transform abundance data to account for huge differences in abundance year to year
summerannuals_2016_sp = sqrt(as.matrix(summerannuals_2016[,c(6:ncol(summerannuals_2016))]))
summerannuals_2017_sp = sqrt(as.matrix(summerannuals_2017[,c(6:ncol(summerannuals_2017))]))
winterannuals_2017_sp = sqrt(as.matrix(winterannuals_2017[,c(6:ncol(winterannuals_2017))]))
summerannuals_sp = sqrt(as.matrix(summerannuals[,c(6:ncol(summerannuals))]))
winterannuals_sp = sqrt(as.matrix(winterannuals[,c(6:ncol(winterannuals))]))

summerannuals_2016_trt = as.factor(summerannuals_2016$treatment)
summerannuals_2017_trt = as.factor(summerannuals_2017$treatment)
winterannuals_2017_trt = as.factor(winterannuals_2017$treatment)
summerannuals_trt = as.factor(summerannuals$treatment)
winterannuals_trt = as.factor(winterannuals$treatment)

summerannuals_year = as.factor(summerannuals$year)
winterannuals_year = as.factor(winterannuals$year)
```

### Summer Annuals 2016
```{r annual_summer2016, echo=FALSE, cache=TRUE}
# cca
sum.ann.2016.cca <- cca(summerannuals_2016_sp ~ summerannuals_2016_trt)

# plotting from Erica's
results=data.frame(scores(sum.ann.2016.cca, display = "sites", scaling = "sites"),
                   treatment = summerannuals_2016_trt)
centroids = aggregate(cbind(CCA1, CA1) ~ treatment, data = results, FUN = mean)

plotobj = ggplot(results, aes(CCA1, CA1)) + 
    stat_ellipse(aes(color = treatment)) +
    geom_point(aes(colour = treatment), data = centroids, size = 3, stroke = 1, shape = 3) +
    geom_point(aes(color = treatment), size=.5) +
    #geom_text(aes(label = plot, color = as.factor(year))) +
    #coord_equal() +                     # biplots only work with equal scaling
    scale_shape_discrete(guide=F) +
    theme(legend.title=element_blank(), legend.position = "bottom",
          legend.key.height = unit(.5, "cm"), legend.spacing = unit(1, "cm"),
          legend.key.width = unit(.5, "cm"), legend.text = element_text(size=8),
          plot.title = element_blank(),
          #plot.title = element_text(size=11),
          axis.title = element_text(size=10)) +
    #scale_color_discrete(direction=-1) +
    scale_colour_manual(values = cbbPalette,
                        breaks=c("control","exclosure"),
                        labels=c("Control", "KR Exclsoure"))

plotobj

# check variance inflation factor (>10 is bad)
print("Variance Inflation Factor (<10 is fine)")
print(vif.cca(sum.ann.2016.cca))
# proportion of variance explained
print("Variance explained:")
print(sum.ann.2016.cca$CCA$tot.chi/sum.ann.2016.cca$tot.chi)
# significance test for individual predictors (type 3 test) (Not Applicable here because I only have one predictor)
print(anova(sum.ann.2016.cca, by="margin"))
# significance test for entire model
#anova(win.cca)
```

### Winter Annuals 2017
```{r annual_winter2017, echo=FALSE, cache=TRUE}
# cca
wint.ann.2017.cca <- cca(winterannuals_2017_sp ~ winterannuals_2017_trt)

# plotting from Erica's
results=data.frame(scores(wint.ann.2017.cca, display = "sites", scaling = "sites"),
                   treatment = winterannuals_2017_trt)
centroids = aggregate(cbind(CCA1, CA1) ~ treatment, data = results, FUN = mean)

plotobj = ggplot(results, aes(CCA1, CA1)) + 
    stat_ellipse(aes(color = treatment)) +
    geom_point(aes(colour = treatment), data = centroids, size = 3, stroke = 1, shape = 3) +
    geom_point(aes(color = treatment), size=.5) +
    #geom_text(aes(label = plot, color = as.factor(year))) +
    #coord_equal() +                     # biplots only work with equal scaling
    scale_shape_discrete(guide=F) +
    theme(legend.title=element_blank(), legend.position = "bottom",
          legend.key.height = unit(.5, "cm"), legend.spacing = unit(1, "cm"),
          legend.key.width = unit(.5, "cm"), legend.text = element_text(size=8),
          plot.title = element_blank(),
          #plot.title = element_text(size=11),
          axis.title = element_text(size=10)) +
    #scale_color_discrete(direction=-1) +
    scale_colour_manual(values = cbbPalette,
                        breaks=c("control","exclosure"),
                        labels=c("Control", "KR Exclsoure"))

plotobj

# check variance inflation factor (>10 is bad)
print("Variance Inflation Factor (<10 is fine)")
print(vif.cca(wint.ann.2017.cca))
# proportion of variance explained
print("Variance explained:")
print(wint.ann.2017.cca$CCA$tot.chi/wint.ann.2017.cca$tot.chi)
# significance test for individual predictors (type 3 test) (Not Applicable here because I only have one predictor)
print(anova(wint.ann.2017.cca, by="margin"))
# significance test for entire model
#anova(win.cca)
```

### Summer Annuals 2017
```{r annual_summer2017, echo=FALSE, cache=TRUE}
# cca
sum.ann.2017.cca <- cca(summerannuals_2017_sp ~ summerannuals_2017_trt)

# plotting from Erica's
results=data.frame(scores(sum.ann.2017.cca, display = "sites", scaling = "sites"),
                   treatment = summerannuals_2017_trt)
centroids = aggregate(cbind(CCA1, CA1) ~ treatment, data = results, FUN = mean)

plotobj = ggplot(results, aes(CCA1, CA1)) + 
    stat_ellipse(aes(color = treatment)) +
    geom_point(aes(colour = treatment), data = centroids, size = 3, stroke = 1, shape = 3) +
    geom_point(aes(color = treatment), size=.5) +
    #geom_text(aes(label = plot, color = as.factor(year))) +
    #coord_equal() +                     # biplots only work with equal scaling
    scale_shape_discrete(guide=F) +
    theme(legend.title=element_blank(), legend.position = "bottom",
          legend.key.height = unit(.5, "cm"), legend.spacing = unit(1, "cm"),
          legend.key.width = unit(.5, "cm"), legend.text = element_text(size=8),
          plot.title = element_blank(),
          #plot.title = element_text(size=11),
          axis.title = element_text(size=10)) +
    #scale_color_discrete(direction=-1) +
    scale_colour_manual(values = cbbPalette,
                        breaks=c("control","exclosure"),
                        labels=c("Control", "KR Exclsoure"))

plotobj

# check variance inflation factor (>10 is bad)
print("Variance Inflation Factor (<10 is fine)")
print(vif.cca(sum.ann.2017.cca))
# proportion of variance explained
print("Variance explained:")
print(sum.ann.2017.cca$CCA$tot.chi/sum.ann.2017.cca$tot.chi)
# significance test for individual predictors (type 3 test) (Not Applicable here because I only have one predictor)
print(anova(sum.ann.2017.cca, by="margin"))
# significance test for entire model
#anova(win.cca)
```

```{r annual_summer_since_flip, echo=FALSE, cache=TRUE}
# # cca
# sum.ann.cca <- cca(summerannuals_sp ~ summerannuals_trt + Condition(summerannuals_year))
# 
# # plotting from Erica's
# results=data.frame(scores(wint.ann.2017.cca, display = "sites", scaling = "sites"),
#                    treatment = summerannuals_2016_trt)
# centroids = aggregate(cbind(CCA1, CA1) ~ treatment, data = results, FUN = mean)
# 
# plotobj = ggplot(results, aes(CCA1, CA1)) + 
#     stat_ellipse(aes(color = treatment)) +
#     geom_point(aes(colour = treatment), data = centroids, size = 3, stroke = 1, shape = 3) +
#     geom_point(aes(color = treatment), size=.5) +
#     #geom_text(aes(label = plot, color = as.factor(year))) +
#     #coord_equal() +                     # biplots only work with equal scaling
#     scale_shape_discrete(guide=F) +
#     theme(legend.title=element_blank(), legend.position = "bottom",
#           legend.key.height = unit(.5, "cm"), legend.spacing = unit(1, "cm"),
#           legend.key.width = unit(.5, "cm"), legend.text = element_text(size=8),
#           plot.title = element_blank(),
#           #plot.title = element_text(size=11),
#           axis.title = element_text(size=10)) +
#     #scale_color_discrete(direction=-1) +
#     scale_colour_manual(values = cbbPalette,
#                         breaks=c("control","exclosure"),
#                         labels=c("Control", "KR Exclsoure"))
# 
# plotobj
# 
# # check variance inflation factor (>10 is bad)
# print("Variance Inflation Factor (<10 is fine)")
# print(vif.cca(wint.ann.2017.cca))
# # proportion of variance explained
# print("Variance explained:")
# print(wint.ann.2017.cca$CCA$tot.chi/wint.ann.2017.cca$tot.chi)
# # significance test for individual predictors (type 3 test) (Not Applicable here because I only have one predictor)
# print(anova(wint.ann.2017.cca, by="margin"))
# # significance test for entire model
# #anova(win.cca)
```

## All Plants by Sampling Period and Treatment

```{r all_plants, echo=FALSE, include=FALSE, cache=TRUE}
summer_2016 <- portalr::plant_abundance(level = 'plot',
                                        shape = 'crosstab') %>% 
  filter(year == '2016',
         season == 'summer',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21))

winter_2017 <- portalr::plant_abundance(level = 'plot',
                                        shape = 'crosstab') %>% 
  filter(year == '2017',
         season == 'winter',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21))

summer_2017 <- portalr::plant_abundance(level = 'plot',
                                        shape = 'crosstab') %>% 
  filter(year == '2017',
         season == 'summer',
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21))

all <- portalr::plant_abundance(level = 'plot',
                                shape = 'crosstab') %>% 
  filter(year > 2014,
         plot %in% c(3, 4, 11, 14, 15, 17, 19, 20, 21))
```

```{r all_plants_cca_prep, echo=FALSE, include=FALSE}
# sqrt transform abundance data to account for huge differences in abundance year to year
summer_2016_sp = sqrt(as.matrix(summer_2016[,c(6:ncol(summer_2016))]))
summer_2017_sp = sqrt(as.matrix(summer_2017[,c(6:ncol(summer_2017))]))
winter_2017_sp = sqrt(as.matrix(winter_2017[,c(6:ncol(winter_2017))]))
all_sp = sqrt(as.matrix(all[,c(6:ncol(all))]))

summer_2016_trt = as.factor(summer_2016$treatment)
summer_2017_trt = as.factor(summer_2017$treatment)
winter_2017_trt = as.factor(winter_2017$treatment)

all_trt = as.factor(all$treatment)
all_year = as.factor(all$year)
```

### Summer All Plants 2016
```{r all_summer2016, echo=FALSE, cache=TRUE}
# cca
sum.all.2016.cca <- cca(summer_2016_sp ~ summer_2016_trt)

# plotting from Erica's
results=data.frame(scores(sum.all.2016.cca, display = "sites", scaling = "sites"),
                   treatment = summer_2016_trt)
centroids = aggregate(cbind(CCA1, CA1) ~ treatment, data = results, FUN = mean)

plotobj = ggplot(results, aes(CCA1, CA1)) + 
    stat_ellipse(aes(color = treatment)) +
    geom_point(aes(colour = treatment), data = centroids, size = 3, stroke = 1, shape = 3) +
    geom_point(aes(color = treatment), size=.5) +
    #geom_text(aes(label = plot, color = as.factor(year))) +
    #coord_equal() +                     # biplots only work with equal scaling
    scale_shape_discrete(guide=F) +
    theme(legend.title=element_blank(), legend.position = "bottom",
          legend.key.height = unit(.5, "cm"), legend.spacing = unit(1, "cm"),
          legend.key.width = unit(.5, "cm"), legend.text = element_text(size=8),
          plot.title = element_blank(),
          #plot.title = element_text(size=11),
          axis.title = element_text(size=10)) +
    #scale_color_discrete(direction=-1) +
    scale_colour_manual(values = cbbPalette,
                        breaks=c("control","exclosure"),
                        labels=c("Control", "KR Exclsoure"))

plotobj

# check variance inflation factor (>10 is bad)
print("Variance Inflation Factor (<10 is fine)")
print(vif.cca(sum.all.2016.cca))
# proportion of variance explained
print("Variance explained:")
print(sum.all.2016.cca$CCA$tot.chi/sum.all.2016.cca$tot.chi)
# significance test for individual predictors (type 3 test) (Not Applicable here because I only have one predictor)
print(anova(sum.all.2016.cca, by="margin"))
# significance test for entire model
#anova(win.cca)
```

### Winter All Plants 2017
```{r all_winter2017, echo=FALSE, cache=TRUE}
# cca
wint.all.2017.cca <- cca(winter_2017_sp ~ winter_2017_trt)

# plotting from Erica's
results=data.frame(scores(wint.all.2017.cca, display = "sites", scaling = "sites"),
                   treatment = winter_2017_trt)
centroids = aggregate(cbind(CCA1, CA1) ~ treatment, data = results, FUN = mean)

plotobj = ggplot(results, aes(CCA1, CA1)) + 
    stat_ellipse(aes(color = treatment)) +
    geom_point(aes(colour = treatment), data = centroids, size = 3, stroke = 1, shape = 3) +
    geom_point(aes(color = treatment), size=.5) +
    #geom_text(aes(label = plot, color = as.factor(year))) +
    #coord_equal() +                     # biplots only work with equal scaling
    scale_shape_discrete(guide=F) +
    theme(legend.title=element_blank(), legend.position = "bottom",
          legend.key.height = unit(.5, "cm"), legend.spacing = unit(1, "cm"),
          legend.key.width = unit(.5, "cm"), legend.text = element_text(size=8),
          plot.title = element_blank(),
          #plot.title = element_text(size=11),
          axis.title = element_text(size=10)) +
    #scale_color_discrete(direction=-1) +
    scale_colour_manual(values = cbbPalette,
                        breaks=c("control","exclosure"),
                        labels=c("Control", "KR Exclsoure"))

plotobj

# check variance inflation factor (>10 is bad)
print("Variance Inflation Factor (<10 is fine)")
print(vif.cca(wint.all.2017.cca))
# proportion of variance explained
print("Variance explained:")
print(wint.all.2017.cca$CCA$tot.chi/wint.all.2017.cca$tot.chi)
# significance test for individual predictors (type 3 test) (Not Applicable here because I only have one predictor)
print(anova(wint.all.2017.cca, by="margin"))
# significance test for entire model
#anova(win.cca)
```

### Summer All Plants 2017
```{r all_summer2017, echo=FALSE, cache=TRUE}
# cca
sum.all.2017.cca <- cca(summer_2017_sp ~ summer_2017_trt)

# plotting from Erica's
results=data.frame(scores(sum.all.2017.cca, display = "sites", scaling = "sites"),
                   treatment = summer_2017_trt)
centroids = aggregate(cbind(CCA1, CA1) ~ treatment, data = results, FUN = mean)

plotobj = ggplot(results, aes(CCA1, CA1)) + 
    stat_ellipse(aes(color = treatment)) +
    geom_point(aes(colour = treatment), data = centroids, size = 3, stroke = 1, shape = 3) +
    geom_point(aes(color = treatment), size=.5) +
    #geom_text(aes(label = plot, color = as.factor(year))) +
    #coord_equal() +                     # biplots only work with equal scaling
    scale_shape_discrete(guide=F) +
    theme(legend.title=element_blank(), legend.position = "bottom",
          legend.key.height = unit(.5, "cm"), legend.spacing = unit(1, "cm"),
          legend.key.width = unit(.5, "cm"), legend.text = element_text(size=8),
          plot.title = element_blank(),
          #plot.title = element_text(size=11),
          axis.title = element_text(size=10)) +
    #scale_color_discrete(direction=-1) +
    scale_colour_manual(values = cbbPalette,
                        breaks=c("control","exclosure"),
                        labels=c("Control", "KR Exclsoure"))

plotobj

# check variance inflation factor (>10 is bad)
print("Variance Inflation Factor (<10 is fine)")
print(vif.cca(sum.all.2017.cca))
# proportion of variance explained
print("Variance explained:")
print(sum.all.2017.cca$CCA$tot.chi/sum.all.2017.cca$tot.chi)
# significance test for individual predictors (type 3 test) (Not Applicable here because I only have one predictor)
print(anova(sum.all.2017.cca, by="margin"))
# significance test for entire model
#anova(win.cca)
```